% introducao
\chapter{Introduction}
 This chapter will drive the reader through the context and motivation of this work followed by the objectives and later the organization of this text is presented.  
 

\section{Context}

% stating how important quality is % 
One of the pillars of sustainable software development is its quality which can basically be defined as internal and external. Internal quality focuses on how software meets its specification and works accordingly to its requirements. External quality is aimed on how well the software is structured and designed. To measure external quality there is the need to execute the software\footnote{Also known as dynamic analysis} either by an end user accessing the system or an automated process like for example functional testing or performance testing. Internal quality however can be verified by either \emph{statical analysis} that is mainly the inspection of the source code itself or by dynamic analysis which means executing the software like for example automated \emph{whitebox testing}\footnote{whitebox testing is the detailed investigation of internal logic and structure of the code \citep{Khan 2012}}.   

With good software quality in mind we take applications to another level where maintainability is increased, correctness is enhanced, defects are identified in early development stages, which can lead up to 100 times reduced costs \citep{Beohm 2001}, and also other characteristics like reusability, reliability and portability are benefited by higher software quality.  

% OSGi and its importance %
A well known and successful way to structure software architecture is to modularize its components allowing easier evolution of the system because smaller decoupled modules are typically easier to maintain than classical applications. In the Java ecosystem there is a moving to modularize the JDK and Java applications with the project Jigsaw \citep{Krill 2012} and also a recent interest in \emph{microservices} \citep{Knorr 2014} arise. Although all this interest in modular application today the only practical working and well known solution for modular Java applications is OSGi \citep{Hall 2011}, a very popular component-based and service-oriented framework for building Java modular applications. OSGi is the \emph{de facto} standard solution for this kind of software since early 2000's and have being used as basis of most JavaEE\footnote{A Java platform dedicated for enterprise applications which are usually secure and robust systems that display, manipulate and store large amounts of complex data maintained by an organization} application servers\footnote{Java application servers are like an extended virtual machine for running applications, transparently handling connections to the database, connections to the Web client, managing components like Enterprise Java Beans(EJB) and so on}, the open source IDE Eclipse\citep{eclipse 2006}, Atlassian Jira and Confluence to cite a few big players using OSGi. 

In the context of software quality and Java modular applications using OSGi there is no known standard way neither well known tools to measure OSGi projects \textit{internal quality} \citep{Hamza 2013} \citep{Wang 2012}. For \emph{external quality} the classical approaches like automated testing are sufficient because in this kind of quality aims in the \emph{behavior} and not the \emph{design} so technology and architecture is usually not taken into account.

        
\section{Objectives}
The main objective of this work is to create a tool to extract software metrics and measure internal quality of OSGi projects where these metrics must reflect good practices in the OSGi world. The main difference the proposed metrics have compared to classical software metrics is that the first will be based on modularity attributes that only exists in modular applications. The tool applies and validate the metrics on real OSGi projects and finally the resulting qualities are analyzed.  

 
\section{Organization}

This text is organized in the following way. First chapter defines the context, motivation and objectives of this work. The second chapter introduces the main concepts and technologies used in this work and is divided into two main sections where the first is focused in the area of software quality like quality measurement, quality metrics, program analysis and quality analysis tools. The second section of chapter two presents Java and OSGi, how standard Java and OSGi are different in respect to quality metrics and why we need different metrics for OSGi. The third chapter presents \textbf{Intrabundle}, an OSGi code introspection tool to measure internal quality, it shows how Intrabundle works, what kind of information it extracts and what metrics it is applying. The fourth chapter analyzes the results Intrabundle produces and validates them to decide if this work has a valid contribution or not. The last chapter presents the conclusions and future work on this subject.